---
title: "Regression Models Assignment - Motor Trend"
author: "R3M79"
date: "29 de Janeiro de 2018"
output: pdf_document
---

# Synopsis

In this project we assume that we're performing an analysis for the magazine Motor Trend, where we'll explore the relation between MPG (miles per galon) and different predictors

# Motor Trend Analysis

## Overview

Motor Trend intends to analyze the relation between MPG (miles per galon) and different variables.
The goal is to answer the following questions:

1. Is an automatic or manual transmission better for MPG

2. Quantify the MPG difference between automatic and manual transmissions

## Exploratory Analysis

```{r setup, include=FALSE }
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, include = FALSE)


```

First we'll load the required libraries and data (mtcars)

```{r loadLibrariesData,message=FALSE, warning=FALSE}

#load libraries
library(dplyr)
library(ggplot2)

#load data
data(mtcars)

#Define Variables for simulation
set.seed(100) # set the seed value for reproducibility

```

### Data Detail and Summary
```{r plotDataPrep,message=FALSE, warning=FALSE}
#prepare data
mydata<-mtcars %>%
    mutate(transmission = ifelse(am=="0","A","M"))%>%
    mutate(am=NULL)

#prepare plot data
plotdata <- ggplot(aes(x=as.factor(transmission), y=mpg,fill=transmission), data=mydata) + geom_boxplot()
plotdata <- plotdata + labs(title = "Automatic vs Manual Transmission")
plotdata <- plotdata + xlab("Transmission")
plotdata <- plotdata + ylab("MPG")

```

Data can be visualized in the boxplot on Appendix - Figures

```{r summDataDisplay,message=FALSE, warning=FALSE,include=TRUE}

#Data Summary
head(mydata)


```
From the boxplot we can see a significant difference between manual and automatic transmission, where we see a higher MPG for manual.

Let's see the Linear model for the outcome MPG with Transmission as the predictor

```{r linearModel1,message=FALSE, warning=FALSE,include=TRUE}

#Linear Model MPG ~ transmission
lmd1<-lm(mpg~transmission,mydata)
summary(lmd1)$coeff
summary(lmd1)$adj.r.squared

```
We can see from the model that there's a significante increase for in MPG for the manual transmission.
Since the value of the adjusted R for model 1 is of only 33,85%, we shoudl consider other variables that may produce better model

## Model Analysis and Selection

Let's first check the correlation between all variables for the mtcars data (plot can be seen on Appendix - Figures)

From the correlation plot it seems that variable cyl, disp, hp and wt have the strongest correlation.

We'll create two new models to verify this.

1. Model with predictors transmission + cyl + disp + hp + wt

```{r linearModel2,message=FALSE, warning=FALSE,include=TRUE}

#Linear Model MPG ~ transmission + cyl + disp + hp + wt
lmd2<-lm(mpg~transmission + cyl + disp + hp + wt,mydata)
summary(lmd2)$coeff
summary(lmd2)$adj.r.squared

```

2. Model with all predictors

```{r linearModel3,message=FALSE, warning=FALSE,include=TRUE}

#Linear Model MPG ~ All variables
lmd3<-lm(mpg~.,mydata)
summary(lmd3)$coeff
summary(lmd3)$adj.r.squared

```

From the values for adjusted R in both new models it looks as the second model is the best, with 82% variance explained

Let's confirm with anova.

```{r modelComp,message=FALSE, warning=FALSE,include=TRUE}

#lets now compare the 3 models
anova(lmd1,lmd2,lmd3)


```
## Conclusion

We clearly see that the second model is the best (MPG ~ Transmission + Cyl + Disp + HP + WT).
Based on this model the difference between Auto and Manual Transmission is of 1.55 MPG (on Appendix - figures we can see the residuals for this model)


# Appendix

##Figures

###BoxPlot for data Mtcars
```{r dataPlot,message=FALSE, warning=FALSE,include=TRUE}

#plot data
plotdata

```

###Data Correlation plot for data Mtcars
```{r dataCorr,message=FALSE, warning=FALSE,include=TRUE}

#Display Data Correlation
pairs(mpg~ . ,data=mtcars)

```

### Residuals for Model 2: MPG ~ Transmission + Cyl + Disp + HP + WT
```{r residuals,message=FALSE, warning=FALSE,include=TRUE}

#display residuals
par(mfrow=c(2,2))
plot(lmd2)

```